---
title: "Task 4: Naples Reef (2010) Purple Urchins"
author: "Chase Tarr"
date: "10/29/2021"
output: html_document
---
# Overview
 - In this document, we will be analyzing data of Purple Urchins located in Naples Reef observed in the year 2010.
 
```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(kableExtra)
library(DT)

urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```
### Comparing mean size of purple urchins in CONTROL

```{r}
purple_urchins_compare <- urchins %>% # pulling data from urchins dataset
  filter(site == "NAPL", # filtering for data only from Naples Reef 2010
         common_name == "Purple Urchin") %>% # filtering data to only pull values from Purple Urchins
  mutate(date = mdy(date), # putting data into correct date format
         year = year(date)) %>% 
  filter(year == "2010")
```

```{r}
ggplot(data = purple_urchins_compare, aes(sample = size, fill = treatment)) +
  geom_qq() +
  facet_wrap(~ treatment) +
  theme_bw() +
  labs(y = "Urchin Size")
 
```
 
This plot represents the size of **Purple Urchins** separated by *treatment*. The *ANNUAL* control shows a slightly bigger urchin size with a normal distribution in both treatments. There is an evident pattern of urchins increasing by incriments of half a cm. 
 
### Creating a table to visualize mean size per treatment
```{r}
purple_urchins_mean <- purple_urchins_compare %>% 
  group_by(treatment, year, common_name) %>%  # grouping by treatment: Either Control or Annual, and including the column 'common_name" for future reference to what we were visualizing
  summarize(
    mean_cm = round(mean(size), 2), # rounding data
    sd_cm = round(sd(size), 2),
    median_cm = median(size),
    samplesize_cm = n()
  )

```
 
## Purple Urchins Mean Size Data Table 
```{r}
datatable(purple_urchins_mean)
```

 This table indicates the *mean*, *standard deviation*, *median*, and *sample size* in cm of the Purple Urchins in Naples Reef. 